Sample-Based Determination of Angular Observables

### 2014-06-24 draft.tex
Tweak app E a little bit

\section{On the Partial Wave Expansions of Interference Terms}
\label{app:partial-waves}

Let $P \equiv P(\vec{\nu},\theta)$ be an observable of only one angle $\theta$ and further non-angular variables $\vec{\nu}$. Further, let
$P$ have an expansion in terms of partial waves $l_1, l_2 = 0,1,2,\dots \hat{=} S,P,D,\dots$ of the underlying amplitudes $A_1$ and $A_2$,
\begin{equation}
    P(\theta) \equiv A_1(\theta) A_2^*(\theta) \equiv \left(\sum_{l_1=0}^\infty A_1^{(l_1)} p_{l_1}^{(m_1)}(\cos\theta)\right) \left(\sum_{l_2=0}^\infty A_1^{(l_2)} p_{l_2}^{(m_2)}(\cos\theta)\right)\,.
\end{equation}
Here $m_1$ and $m_2$ denote the third component of the angular momentum of the respective amplitude $A_1$ or $A_2$, and we impose $l_1 \geq l_2$.\\

From positivity of the amplitudes one immediately finds
\begin{equation}
    \int \dd\cos\theta |A_i(\theta)|^2 = \sum_{l_i = 0}^\infty |A_i^{(l_i)}|^2 n_{l_i, m_i} < \infty\,,
\end{equation}
where we introduce $n_{l,m}$ via the scalar product of two associated Legendre polynomials,
\begin{equation}
    \label{eq:legendre-scalar-product}
    n_{l, m} \delta_{l, l'} \equiv \int_{0}^\pi p_{l}^{(m)}(\cos\theta) p_{l'}^{(m)} (\cos\theta) \dd \cos\theta = \frac{2}{(2 l + 1)} \frac{(l + m)!}{(l - m)!} \delta_{l, l'}\,.
\end{equation}
Using the large-$l$ behaviour of $n_{l_i,m_i}$ one finds that with increasing $l_i$ the partial waves amplitudes must asymptotically fall off,
\begin{equation}
    |A_i^{(l_i)}(\cos\theta)| \Big|_{l_i \gg m_i} \sim \frac{1}{\sqrt{l_i}}\,.
\end{equation}
This motivates to cut-off of the partial-wave expansion at some angular momentum $L$. Will will show in the following that such a cut-off
is compatible with defining a basis of angular observables as coefficients of Legendre polynomials in $\cos\theta$.\\

We propose the decomposition of $P$ in terms of the associated Legendre polynomials $p_{j}^{(m)}(\cos\theta)$, with $m=m_1 + m_2$, and
$l_1 - l_2 \leq j \leq l_1 + l_2$,
\begin{equation}
    P(\vec{\nu}, \theta) = \sum_j P_{j,m}(\vec{\nu}) p_{j}^{(m)}(\cos\theta)\,.
\end{equation}
This parametrization has two merits. First, we can immediately project out the angular observables $P_{j,m}$ by means of \refeq{legendre-scalar-product}:
\begin{equation}
    P_{j,m}(\vec{\nu}) = \frac{1}{n_{j,m}} \int_{-1}^{+1} P(\vec{\nu},\cos \theta) p_{j}^{(m)}(\cos\theta) \dd\cos\theta\,.
\end{equation}
Second, we can immediately express $P_{j,m}(\vec{\nu})$ in terms of the partial-wave amplitudes,
\begin{equation}
    \label{eq:partial-wave-observable-infinite}
    \begin{aligned}
        P_{j,m}(\vec{\nu})
            & = \frac{1}{n_{j,m}} \int_{-1}^{+1} p_{j}^{(m_1 + m_2)}(\cos\theta) \sum_{l_1,l_2=0}^\infty A_{1}^{l_1} p_{l_1}^{(m_1)}(\cos\theta) \big(A_{2}^{(l_2)}\big)^* p_{l_2}^{(m_2)}(\cos\theta)\\
            & = \sum_{l_1,l_2=0}^\infty A_{1}^{l_1} \big(A_{2}^{(l_2)}\big)^* \frac{T_{l_1,l_2,j}^{(m_1,m_2)}}{n_{j,m}}\,,\qquad\text{with }m = m_1 + m_2\,.
    \end{aligned}
\end{equation}
In the last step, we use Gaunt's formula \cite{Gaunt:1929} to integrate a triple product of associated Legendre polynomials,
\begin{equation}
\begin{aligned}
    T_{l_1,l_2,j}^{(m_1,m_2)}
        & = \int_{-1}^{+1} p_{j}^{(m_1 + m_2)}(\cos\theta) p_{l_1}^{(m_1)}(\cos\theta) p_{l_2}^{(m_2)}(\cos\theta) \dd\cos\theta\\
        & = (-1)^{s - l_1 - m_2} \frac{2 (l_1 + m_1)! (l_2 + m_2)! (2s - 2 l_2)! s!}{(l_1 - m_1)! (s - j)! (s - l_1)! (s - l_2)! (2s + 1)!}\\
        & \times \sum_{t=p}^q (-1)^t \frac{(j + m + t)!(l_1 + l_2 - m - t)!}{t! (j - m - t)! (l_1 - l_2 + m + t)! (l_2 - m_2 - t)!}\,,
\end{aligned}
\end{equation}
where
\begin{equation}
\begin{aligned}
    m & = m_1 + m_2\,, &
    m_1 & \geq m_2\,,  \\
    j, l_1, l_2 & \geq 0\,, &
    m, m_1, m_2 & \geq 0\,,
\end{aligned}
\end{equation}
and
\begin{equation}
\begin{aligned}
    s & = \frac{j + l_1 + l_2}{2}\,, &
    p & = \max(0, l_2 - l_1 - m)\,, &
    q & = \min(l_1 + l_2 - m, j - m, l_2 - m_2)\,.
\end{aligned}
\end{equation}
The necessary condition for $T \neq 0$ are
\begin{equation}
    s \in \mathbb{N}\qquad \wedge \qquad l_1 - l_2 \leq l \leq l_1 + l_2,.
\end{equation}
The latter condition is well known from adding angular momenta.

>>>>

\section{On the Partial Wave Expansions of Angular Observables}
\label{app:partial-waves}

Let $P \equiv P(\vec{\nu},\theta)$ be an observable of only one angle $\theta$ and further non-angular variables $\vec{\nu}$. Further, let
$P$ have an expansion in terms of partial waves $l_1, l_2 = 0,1,2,\dots \hat{=} S,P,D,\dots$ of the underlying amplitudes $A_1$ and $A_2$,
\begin{equation}
    \label{eq:def-partial-wave-observable}
    P(\theta) \equiv F\big(A_1(\theta) A_2^*(\theta)\big) \equiv \left(\sum_{l_1=0}^\infty A_1^{(l_1)} p_{l_1}^{(m_1)}(\cos\theta)\right) \left(\sum_{l_2=0}^\infty A_1^{(l_2)} p_{l_2}^{(m_2)}(\cos\theta)\right)\,,\qquad F=\Re{},\Im{}\,.
\end{equation}
Here $m_1$ and $m_2$ denote the third component of the angular momentum of the respective amplitude $A_1$ or $A_2$, and we impose $m_1 \geq m_2$.\\

From positivity of the amplitudes one immediately finds
\begin{equation}
    \int \dd\cos\theta |A_i(\theta)|^2 = \sum_{l_i = 0}^\infty |A_i^{(l_i)}|^2 n_{l_i, m_i} < \infty\,,
\end{equation}
where we introduce $n_{l,m}$ via the scalar product of two associated Legendre polynomials,
\begin{equation}
    \label{eq:legendre-scalar-product}
    n_{l, m} \delta_{l, l'} \equiv \int_{0}^\pi p_{l}^{(m)}(\cos\theta) p_{l'}^{(m)} (\cos\theta) \dd \cos\theta = \frac{2}{(2 l + 1)} \frac{(l + m)!}{(l - m)!} \delta_{l, l'}\,.
\end{equation}
Using the large-$l$ behaviour of $n_{l,m}$ one finds that with increasing $l_i$ the partial waves amplitudes must asymptotically fall off,
\begin{equation}
    |A_i^{(l_i)}(\cos\theta)| \Big|_{l_i \gg m_i} \sim \frac{1}{\sqrt{l_i}}\,.
\end{equation}
This motivates to cut the partial-wave expansion off at some angular momentum $L$.\\

Will will show in the following that such a cut off
is compatible with defining a basis of angular observables as coefficients of Legendre polynomials in $\cos\theta$. Since this expansion implies a well-defined total angular momentum for each observables, one ensures that the observables can in fact be disentangled experimentally.\\

We propose the decomposition of $P$ in terms of the associated Legendre polynomials $p_{j}^{(m)}(\cos\theta)$, with total angular momentum $j$ and its third component $m=m_1 + m_2$.
\begin{equation}
    P(\vec{\nu}, \theta) = \sum_j P_{j,m}(\vec{\nu}) p_{j}^{(m)}(\cos\theta)\,.
\end{equation}
This parametrization has two merits. First, we can immediately project out the angular observables $P_{j,m}$ by means of \refeq{legendre-scalar-product}:
\begin{equation}
    P_{j,m}(\vec{\nu}) = \frac{1}{n_{j,m}} \int_{-1}^{+1} P(\vec{\nu},\cos \theta) p_{j}^{(m)}(\cos\theta) \dd\cos\theta\,.
\end{equation}
Second, we can immediately express $P_{j,m}(\vec{\nu})$ in terms of the partial-wave amplitudes,
\begin{equation}
    \label{eq:partial-wave-observable-infinite}
    \begin{aligned}
        P_{j,m}(\vec{\nu})
            & = \frac{1}{n_{j,m}} \int_{-1}^{+1} p_{j}^{(m_1 + m_2)}(\cos\theta) \sum_{l_1,l_2=0}^\infty F\left(A_{1}^{l_1} p_{l_1}^{(m_1)}(\cos\theta) \big(A_{2}^{(l_2)}\big)^*\right) p_{l_2}^{(m_2)}(\cos\theta)\\
            & = \sum_{l_1,l_2=0}^\infty F\left(A_{1}^{l_1} \big(A_{2}^{(l_2)}\big)^*\right) \frac{T_{l_1,l_2,j}^{(m_1,m_2)}}{n_{j,m}}\,,\qquad\text{with }m = m_1 + m_2\,.
    \end{aligned}
\end{equation}
In the last step, we use Gaunt's formula \cite{Gaunt:1929} to integrate a triple product of associated Legendre polynomials,
\begin{equation}
\begin{aligned}
    T_{l_1,l_2,j}^{(m_1,m_2)}
        & = \int_{-1}^{+1} p_{j}^{(m_1 + m_2)}(\cos\theta) p_{l_1}^{(m_1)}(\cos\theta) p_{l_2}^{(m_2)}(\cos\theta) \dd\cos\theta\\
        & = (-1)^{s - l_1 - m_2} \frac{2 (l_1 + m_1)! (l_2 + m_2)! (2s - 2 l_2)! s!}{(l_1 - m_1)! (s - j)! (s - l_1)! (s - l_2)! (2s + 1)!}\\
        & \times \sum_{t=p}^q (-1)^t \frac{(j + m + t)!(l_1 + l_2 - m - t)!}{t! (j - m - t)! (l_1 - l_2 + m + t)! (l_2 - m_2 - t)!}\,,
\end{aligned}
\end{equation}
where
\begin{equation}
\begin{aligned}
    m & = m_1 + m_2\,, &
    m_1 & \geq m_2\,,  \\
    j, l_1, l_2 & \geq 0\,, &
    m, m_1, m_2 & \geq 0\,,
\end{aligned}
\end{equation}
and
\begin{equation}
\begin{aligned}
    s & = \frac{j + l_1 + l_2}{2}\,, &
    p & = \max(0, l_2 - l_1 - m)\,, &
    q & = \min(l_1 + l_2 - m, j - m, l_2 - m_2)\,.
\end{aligned}
\end{equation}
The necessary condition for $T \neq 0$ are
\begin{equation}
    \label{eq:angular-momentum-addition}
    s \in \mathbb{N}\qquad \wedge \qquad l_1 - l_2 \leq l \leq l_1 + l_2,.
\end{equation}
The latter condition is well known from adding angular momenta. Note, however, that
the sum in \refeq{def-partial-wave-observable} goes to infinitely high angular momenta $l_1$ and $l_2$. As a consequence
of this and of \refeq{angular-momentum-addition}, the angular observables $P_{j,m}$
consist of sums with infinite terms. It is then up to theoretical analyses to
estimate or calculate the impact of individual contributions, with or without the cut off $L$.
For such an analysis in the decay $B\to K\pi\ell^+\ell^-$ we refer to \cite{Das:2014}.
